External Patch Group Prior Guided Internal Prior Learning for Real Image Denoising

### 2016-10-26 CVPR2017_Review_Unified.tex
add

Natural images often demonstrate repepitive local patterns, this nonlocal self-similarity (NSS) property is a key successful factor for many image denoising methods \cite{nlm,bm3d,lssc,ncsr,wnnm,pgpd}. In this section, we formulate the Patch Group prior learned on natural color images. Similar to \cite{pgpd}, the patch group (PG) is defined as a group of similar patches to the local patch. The patch group mean is destracted, and hence different groups patches can share similar PGs. In this way, the space natural image patches to be modeled is largely reduced. In this work, each local patch extracted from RGB images is of size $p\times p \times 3$. Then we search the $M$ most similar patches $\{\mathbf{x}_{m}\}_{m=1}^{M}$ around each local patch through Euclidean distance, in a local window of size $W\times W$. The $\mathbf{x}_{m}\in \mathbb{R}^{3p^{2}\times1}$ is a patch vector formed by combining the 3 patch vectors (of size $p^{2}\times 1$) in R, G, B channels. The mean vector of this PG is $\boldsymbol{\upmu}=\frac{1}{M}\sum_{m=1}^{M}\mathbf{x}_{m}$, and the group mean subtracted PG is defined as $\mathbf{\overline{X}}\triangleq \{\mathbf{\overline{x}}_{m}=\mathbf{x}_{m}-\boldsymbol{\upmu}\}, m=1,...,M$. Assume we have extracted $N$ PGs from a set of external natural images, and the $n$-th PG is defined as $\mathbf{\overline{X}}_{n}\triangleq \{\mathbf{\overline{x}}_{n,m}\}_{m=1}^{M}, n=1,...,N$. We employ the Gaussian Mixture Model (GMM) to learn the external patch group based NSS prior. In this model, the likelihood of the $n$-th PG $\{\mathbf{\overline{X}}_{n}\}$ can be calculated as
\vspace{-2mm}
\begin{equation}\label{equ1}\vspace{-2mm}
P(\mathbf{\overline{X}}_{n})  = \sum\nolimits_{k=1}^{K}\pi_{k}\prod\nolimits_{m=1}^{M}\mathcal{N}(\mathbf{\overline{x}}_{n,m}|\boldsymbol{\upmu}_{k},\mathbf{\Sigma}_{k}),
\end{equation}
where $K$ is the number of Gaussians and the parameters $\pi_{k}$, $\boldsymbol{\upmu}_{k}$, $\mathbf{\Sigma}_{k}$ are mixture weight, mean vector, and covariance matrix of the $k$-th Gaussian, respectively. By assuming that all the PGs are independently sampled, the overall objective log-likelihood function is
\vspace{-2mm}
\begin{equation}\label{equ2}\vspace{-2mm}
\begin{split}
\ln\mathcal{L}=\sum_{n=1}^{N} \ln(\sum_{k=1}^{K}\pi_{k}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{x}}_{n,m}|\boldsymbol{\upmu}_{k},\mathbf{\Sigma}_{k})).
\end{split}
\end{equation} 
We maximize the above objective function via EM algorithm \cite{em} and finally obtain the GMM model with learned parameters as the learned external PG prior.

@@

\vspace{-2mm}
\subsubsection{Guided Internal Subspace Selection}

Similar to the previous section, we extract noisy PGs from the given real noisy image and extract the mean substracted PG, which is defined as $\mathbf{\overline{Y}}$. Noted that, the mean vectors of external PGs are discarded in the external prior learning stage while the mean vectors of internal PGs are saved for recovering. For better adaptivity, we project the PG $\mathbf{\overline{Y}}$ into its most suitable Gaussian component (subspace) of the GMM trained by external PGs. The selection can be done by firstly calculating the posterior probability of "$\mathbf{\overline{Y}}$ belonging to the $k$th Gaussian component":
\vspace{-2mm}
\begin{equation}\label{equ3}\vspace{-2mm}
P(k|\mathbf{\overline{Y}})=\frac{\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{k})}{\sum_{l=1}^{K}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{l})},
\end{equation}
and then selecting the component with the maximum A-posteriori (MAP) probability $\ln P(k|\mathbf{\overline{Y}})$ as the most suitable subspace for $\mathbf{\overline{Y}}$.

\vspace{-2mm}
\subsubsection{Guided Internal Orthogonal Dictionary Learning}

To better and adaptively charactering the internal PGs from the testing image, we need learn a more specific dictionary for noisy PGs assigned into each cluster. For notation simplicity, we ignore the index of subspace $k$. The internal PGs $\mathbf{Y}$ form a subspace which can be obtained by singular value decomposition (SVD),
\begin{equation}\label{equ4}
\begin{split}
&\min_{\mathbf{D}_{i}\in \mathbb{R}^{3p^2\times r},\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
The singular vectors capture the statistical structures of NSS variations in natural images, while the singular values in $\mathbf{S}$ represent the significance of these singular vectors. Fig.\ 4 shows the singular vectors for one Gaussian component.

@@

\textbf{Updating Sparse Coefficients}: given the initialization orthogonal dictioanry $\textbf{D}_{i}^{(t)}$, the sparce coefficients 
$\textbf{A}^{(t)}$ are obtained via solving
\begin{equation}\label{equ5}
\textbf{A}^{(t)}:=\argmin_{\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}^{(t)}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}.
\end{equation}
This problem has closed-form solution by $\mathbf{A}^{*}=T_{\lambda}(\mathbf{\hat{D}}^{T}\mathbf{Y})$, 
where $T_{\lambda}(\mathbf{A}) = \text{sgn}(\mathbf{A})\odot\max(\mathbf{A},\lambda)$ is a soft-thresholding function.

\textbf{Updating Orthogonal Dictionary}: given the sparse coefficients $\textbf{A}^{(0)}$, the sparce coefficients 
$\textbf{A}^{(t)}$ are obtained via solving
\begin{equation}\label{equ6} 
\begin{split}
\textbf{D}_{i}^{(t+1)}:=
&
\argmin_{\textbf{D}_{i}\in\mathbb{R}^{3p^2\times r}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}]\mathbf{A}^{(t)}\|_{F}^{2}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
Dividing the sparse coefficients $\mathbf{A}=[\mathbf{A}_{e}^{T}\ \mathbf{A}_{i}^{T}]^{T}$, where $\mathbf{A}_{e}$ and $\mathbf{A}_{i}$ denote the coefficients over external and internal dictionary $\mathbf{D}_{e}$ and $\mathbf{D}_{i}$. According to the Proposition 2.2 in \cite{bao2013fast}, the problem (\ref{equ6}) 
%\begin{equation}
%\begin{split}
%\min_{\textbf{D}_{i}\in\mathbb{R}^{3p^2\times r}}
%&\|\mathbf{Y}-\mathbf{D}_{e}\mathbf{A}_{e}+\mathbf{D}_{i}\mathbf{A}_{i}\|_{F}^{2}
%\\
%&
%\text{s.t.}
%\quad
%\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
%\end{split}
%\end{equation}
has a closed-form solution $\mathbf{D}_{i}^{*}=\mathbf{U}\mathbf{V}^{T}$, where $\mathbf{U}$ and $\mathbf{V}$ are
the orthogonal matrices obtained by the following SVD
\begin{equation}\label{equ7}
(\mathbf{I}-\mathbf{D}_{e}\mathbf{D}_{e}^{T})\mathbf{Y}\mathbf{A}_{i}^{T}
=
\mathbf{U}\Sigma\mathbf{V}^{T}
\end{equation}
With these solutions, the final obtained dictionary $\mathbf{D} = [\mathbf{D}_{e}\ \mathbf{D}_{i}]$ are orthogonal ictionary. This can be proved by the following equation
\begin{equation}\label{equ8}
\mathbf{D}^{T}\mathbf{D} = 
\left(\begin{array}{c}
\mathbf{D}_{e}^{T}
\\
\mathbf{D}_{i}^{T}
\end{array}\right)
(\mathbf{D}_{e}\ \mathbf{D}_{i})
=
\left(\begin{array}{cc}
\mathbf{D}_{e}^{T}\mathbf{D}_{e}\ \mathbf{D}_{e}^{T}\mathbf{D}_{i}
\\
\mathbf{D}_{i}^{T}\mathbf{D}_{e}\ \mathbf{D}_{i}^{T}\mathbf{D}_{i}
\end{array}\right)
=
\mathbf{I}
\end{equation}

\subsection{Discussions on Learning External Priors and Internal Priors}
Until now, we have divided the noisy PGs into multiple internal subspaces. Here we take a deep analysis on how the external NSS prior guide the subspace learning of internal PGs. The help are at least threefold. Firstly, through MAP in (\ref{equ3}), the external prior guides the noisy PGs to be clustered into the correct subspaces. If we cluster the noisy PGs in an automatical way, the subspaces we learned will be highly degraded by the signal dependent noise. Secondly, the guidance of external prior for internal clustering is more efficient than directly clustering the internal noisy PGs. It only needs to calculate the MAP probability via the equation (\ref{equ3}) while the internal clustering via GMM is time-consuming on EM algorithm \cite{em}. Thirdly, due to the correct guidance of external prior, the structual decomposition via SVD of each subspace is more adaptive. This will bring better denoising performance than the methods only using the external information. The \emph{mutual incoherence} $\mu(\mathbf{U})$ \cite{donoho2003optimally}, which is difined as
\begin{equation}\label{equ9}
\mu(\mathbf{U})=\max_{i~=j}\frac{|\mathbf{d}_{i}^{T}\mathbf{d}_{j}|}{\|\mathbf{d}_{i}\|_{2}\|\mathbf{d}_{j}\|_{2}}
\end{equation},
is a measure of quality of dictionary.

>>>>

Natural images often demonstrate repepitive local patterns, this nonlocal self-similarity (NSS) property is a key successful factor for many image denoising methods \cite{nlm,bm3d,lssc,ncsr,wnnm,pgpd}. In this section, we formulate the Patch Group prior learned on natural color images. Similar to \cite{pgpd}, the patch group (PG) is defined as a group of similar patches to the local patch. The patch group mean is destracted, and hence different groups patches can share similar PGs. In this way, the space natural image patches to be modeled is largely reduced. In this work, each local patch extracted from RGB images is of size $p\times p \times 3$. Then we search the $M$ most similar patches $\{\mathbf{x}_{m}\}_{m=1}^{M}$ around each local patch through Euclidean distance, in a local window of size $W\times W$. The $\mathbf{x}_{m}\in \mathbb{R}^{3p^{2}\times1}$ is a patch vector formed by combining the 3 patch vectors (of size $p^{2}\times 1$) in R, G, B channels. The mean vector of this PG is $\boldsymbol{\upmu}=\frac{1}{M}\sum_{m=1}^{M}\mathbf{x}_{m}$, and the group mean subtracted PG is defined as $\mathbf{\overline{X}}\triangleq \{\mathbf{\overline{x}}_{m}=\mathbf{x}_{m}-\boldsymbol{\upmu}\}, m=1,...,M$. Assume we have extracted $N$ PGs from a set of external natural images, and the $n$-th PG is defined as $\mathbf{\overline{X}}_{n}\triangleq \{\mathbf{\overline{x}}_{n,m}\}_{m=1}^{M}, n=1,...,N$. We employ the Gaussian Mixture Model (GMM) to learn the external patch group based NSS prior. In this model, the likelihood of the $n$-th PG $\{\mathbf{\overline{X}}_{n}\}$ can be calculated as
\vspace{-2mm}
\begin{equation}\label{equ1}\vspace{-2mm}
P(\mathbf{\overline{X}}_{n})  = \sum\nolimits_{k=1}^{K}\pi_{k}\prod\nolimits_{m=1}^{M}\mathcal{N}(\mathbf{\overline{x}}_{n,m}|\boldsymbol{\upmu}_{k},\mathbf{\Sigma}_{k}),
\end{equation}
where $K$ is the number of Gaussians and the parameters $\pi_{k}$, $\boldsymbol{\upmu}_{k}$, $\mathbf{\Sigma}_{k}$ are mixture weight, mean vector, and covariance matrix of the $k$-th Gaussian, respectively. By assuming that all the PGs are independently sampled, the overall objective log-likelihood function is
\vspace{-2mm}
\begin{equation}\label{equ2}\vspace{-2mm}
\begin{split}
\ln\mathcal{L}=\sum_{n=1}^{N} \ln(\sum_{k=1}^{K}\pi_{k}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{x}}_{n,m}|\boldsymbol{\upmu}_{k},\mathbf{\Sigma}_{k})).
\end{split}
\end{equation} 
We maximize the above objective function via EM algorithm \cite{em} and finally obtain the GMM model with learned parameters as the learned external PG prior. Similar to \cite{pgpd}, the mean vector of each cluster is natural zeros, i.e., $\boldsymbol{\upmu}_{k}=\mathbf{0}$.

@@

\vspace{-2mm}
\subsubsection{Guided Internal Subspace Selection}
\vspace{-1mm}

Given a real noisy image, we extract the noisy PGs and corresponding mean vectors. Each mean substracted PG is defined as $\mathbf{\overline{Y}}\triangleq \{\mathbf{\overline{x}}_{m}\}_{m=1}^{M}$. Noted that, different from the external PGs, the mean vectors of the internal noisy PGs are saved for recovering. For adaptivity, we project the PG $\mathbf{\overline{Y}}$ into its most suitable Gaussian component (subspace) of the GMM learned on external PGs. The subspace most suitable for $\mathbf{\overline{Y}}$ is selected by firstly calculating the posterior probability of "$\mathbf{\overline{Y}}$ belonging to the $k$th Gaussian component":
\vspace{-3mm}
\begin{equation}\label{equ3}\vspace{-2mm}
P(k|\mathbf{\overline{Y}})=\frac{\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{k})}{\sum_{l=1}^{K}\prod_{m=1}^{M}\mathcal{N}(\mathbf{\overline{y}}_{m}|\mathbf{0},\mathbf{\Sigma}_{l})},
\end{equation}
and then choosing the component with the maximum A-posteriori (MAP) probability $\ln P(k|\mathbf{\overline{Y}})$.

\vspace{-2mm}
\subsubsection{Guided Internal Orthogonal Dictionary Learning}
\vspace{-2mm}

Now each internal noisy PG has been assigned to its most suitable Gaussian component $\mathcal{N}(\mathbf{0},\mathbf{\Sigma}_{k})$ (subspace). In order to recover the latent clean PG, we employ the sparse coding framework under which we need a suitable dictionary for coding. As we know, given a dictionary $\mathbf{D}$, its \emph{mutual incoherence} $\mu(\mathbf{D})$ \cite{donoho2003optimally}, which is difined as
\vspace{-2mm}
\begin{equation}\label{equ4}
\vspace{-2mm}
\mu(\mathbf{D})=\max_{i~=j}\frac{|\mathbf{d}_{i}^{T}\mathbf{d}_{j}|}{\|\mathbf{d}_{i}\|_{2}\|\mathbf{d}_{j}\|_{2}},
\end{equation}
is a measure of quality of dictionary. Since the \emph{mutual incoherence} of a orthogonal dictionary is $0$, its quality is often better than non-orthogonal dictionaries.

external subspace through covariance matrix via singular value decomposition (SVD) 
\begin{equation}\label{equ4}
\mathbf{\Sigma}=\mathbf{U}_{e}\mathbf{S}_{e}\mathbf{V}_{e}^{T}.
\end{equation}
can learn for each subspace an internal orthogonal dictionary adaptive to the noisy PGs. For notation simplicity, we ignore the index of subspace $k$.

\begin{equation}\label{equ5}
\begin{split}
&\min_{\mathbf{D}_{i}\in \mathbb{R}^{3p^2\times r},\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
The singular vectors capture the statistical structures of NSS variations in natural images, while the singular values in $\mathbf{S}$ represent the significance of these singular vectors. Fig.\ 4 shows the singular vectors for one Gaussian component.

@@

\textbf{Updating Sparse Coefficients}: given the initialization orthogonal dictioanry $\textbf{D}_{i}^{(t)}$, the sparce coefficients 
$\textbf{A}^{(t)}$ are obtained via solving
\begin{equation}\label{equ6}
\textbf{A}^{(t)}:=\argmin_{\textbf{A}\in\mathbb{R}^{3p^2\times MN}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}^{(t)}]\mathbf{A}\|_{F}^{2}+\lambda\|\mathbf{A}\|_{1}.
\end{equation}
This problem has closed-form solution by $\mathbf{A}^{*}=T_{\lambda}(\mathbf{\hat{D}}^{T}\mathbf{Y})$, 
where $T_{\lambda}(\mathbf{A}) = \text{sgn}(\mathbf{A})\odot\max(\mathbf{A},\lambda)$ is a soft-thresholding function.

\textbf{Updating Orthogonal Dictionary}: given the sparse coefficients $\textbf{A}^{(0)}$, the sparce coefficients 
$\textbf{A}^{(t)}$ are obtained via solving
\begin{equation}\label{equ7} 
\begin{split}
\textbf{D}_{i}^{(t+1)}:=
&
\argmin_{\textbf{D}_{i}\in\mathbb{R}^{3p^2\times r}}
\|\mathbf{Y}-[\mathbf{D}_{e}\ \mathbf{D}_{i}]\mathbf{A}^{(t)}\|_{F}^{2}
\\
&
\text{s.t.}
\quad
\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
\end{split}
\end{equation}
Dividing the sparse coefficients $\mathbf{A}=[\mathbf{A}_{e}^{T}\ \mathbf{A}_{i}^{T}]^{T}$, where $\mathbf{A}_{e}$ and $\mathbf{A}_{i}$ denote the coefficients over external and internal dictionary $\mathbf{D}_{e}$ and $\mathbf{D}_{i}$. According to the Proposition 2.2 in \cite{bao2013fast}, the problem (\ref{equ6}) 
%\begin{equation}
%\begin{split}
%\min_{\textbf{D}_{i}\in\mathbb{R}^{3p^2\times r}}
%&\|\mathbf{Y}-\mathbf{D}_{e}\mathbf{A}_{e}+\mathbf{D}_{i}\mathbf{A}_{i}\|_{F}^{2}
%\\
%&
%\text{s.t.}
%\quad
%\mathbf{D}_{i}^{T}\mathbf{D}_{i} = \mathbf{I}_{r},\ \mathbf{D}_{e}^{T}\mathbf{D}_{i} = \mathbf{0},
%\end{split}
%\end{equation}
has a closed-form solution $\mathbf{D}_{i}^{*}=\mathbf{U}\mathbf{V}^{T}$, where $\mathbf{U}$ and $\mathbf{V}$ are
the orthogonal matrices obtained by the following SVD
\begin{equation}\label{equ7}
(\mathbf{I}-\mathbf{D}_{e}\mathbf{D}_{e}^{T})\mathbf{Y}\mathbf{A}_{i}^{T}
=
\mathbf{U}\Sigma\mathbf{V}^{T}
\end{equation}
With these solutions, the final obtained dictionary $\mathbf{D} = [\mathbf{D}_{e}\ \mathbf{D}_{i}]$ are orthogonal ictionary. This can be proved by the following equation
\begin{equation}\label{equ8}
\mathbf{D}^{T}\mathbf{D} = 
\left(\begin{array}{c}
\mathbf{D}_{e}^{T}
\\
\mathbf{D}_{i}^{T}
\end{array}\right)
(\mathbf{D}_{e}\ \mathbf{D}_{i})
=
\left(\begin{array}{cc}
\mathbf{D}_{e}^{T}\mathbf{D}_{e}\ \mathbf{D}_{e}^{T}\mathbf{D}_{i}
\\
\mathbf{D}_{i}^{T}\mathbf{D}_{e}\ \mathbf{D}_{i}^{T}\mathbf{D}_{i}
\end{array}\right)
=
\mathbf{I}
\end{equation}

\subsection{Discussions on Learning External Priors and Internal Priors}
Until now, we have divided the noisy PGs into multiple internal subspaces. Here we take a deep analysis on how the external NSS prior guide the subspace learning of internal PGs. The help are at least threefold. Firstly, through MAP in (\ref{equ3}), the external prior guides the noisy PGs to be clustered into the correct subspaces. If we cluster the noisy PGs in an automatical way, the subspaces we learned will be highly degraded by the signal dependent noise. Secondly, the guidance of external prior for internal clustering is more efficient than directly clustering the internal noisy PGs. It only needs to calculate the MAP probability via the equation (\ref{equ3}) while the internal clustering via GMM is time-consuming on EM algorithm \cite{em}. Thirdly, due to the correct guidance of external prior, the structual decomposition via SVD of each subspace is more adaptive. This will bring better denoising performance than the methods only using the external information.
